---
output:
  word_document: default
  html_document: default
---
```{r include=FALSE}

# Package names
packages <- c("tidyverse", "tidymodels", "caret", "gridExtra", "vip", "ranger", "skimr")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if(any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

```

```{r}

# Reading in the data set
drug = read_csv("drug_data-1.csv")

# Renaming the data
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity","Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive","SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis","Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh","LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")

str(drug)

# Changing CL Values
drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"


# Mutating Names
drug_clean = drug %>% 
  mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44","45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =c("Under16", "At16", "At17", "At18", "SomeCollege","ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
  mutate(Country = factor(Country,labels = c("USA", "NewZealand", "Other", "Australia","Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity,labels = c("Black", "Asian", "White", "White/Black", "Other","White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%select(-ID)

# Checking work
str(drug_clean)

```
```{r}
# Dropping unnecessary variables
drug_clean = drug_clean %>% 
  select(!(Alcohol:Mushrooms)) %>% 
  select(!(Semer:VSA))
  names(drug_clean)

```

### Task 1: Check for missing data in our “drug_clean” dataframe. Is there any missingness? If so, identify andimplement a reasonable strategy to deal with the missingness.

```{r}


# Checking for 'missingness'
skim(drug_clean)

```
There appears to be no missing data. Considering we are using Factor and Numeric data, we can be confident in the skimr package's output showing 0 missing data. 

### Task 2: Split the dataset into training (70%) and testing (30%) sets. Use a set.seed of 1234. Stratify by the“Nicotine” variable.

```{r}
# Splitting training and testing sets
set.seed(1234)
drug_split = initial_split(drug_clean, prob = 0.70, strata = Nicotine)
train = training(drug_split)
test = testing(drug_split)

```

### Task 3: Create appropriate visualizations (12 in all) to examine the relationships between each variable and“Nicotine”. Use grid.arrange (from the gridExtra package) to organize these visuals (perhaps in groups of fourvisualizations?). Comment on the relationship between each variable and “Nicotine”.


```{r}

# plotting the Factor Values
p1 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p3 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p4 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))


grid.arrange(p1,p2,p3,p4)

```
**Based on the data:**
* Gender: Men use more Nicotine in their gender than females do
* Education: It appears that though the amount in each category varies - a higher education level tends to lower the nicotine usage
* Country: Ireland has the higher rate of usage in its country and the UK has the lowest usage (highest non-usage)
Ethnicity: It appears that White/Black and White/Asian have higher usage within their own groups. Conversely, Black and Asian groups have the lowest amount of usage (higher non-usage). 

```{r}

# Plotting 1 factor value and the remaining numeric values
p5 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
p6 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p7 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p8 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()


grid.arrange(p5, p6,p7,p8, ncol = 2)

```
**Based on the data:**
* Age: As age increases, there appears to be less usage (higher amounts of non-usage)
* Nscore: Neurotic people tend to use slightly more nicotine than those with lower N-scores
* Escore: Based on the values, it's hard to tell if extroverted people use slightly more or less. It appears to show slightly more toward the "yes" side - though it is close
* Oscore: People with more 'openness to experience' show higher amounts of usage - which makes sense.

```{r}

# Plotting remaining numeric values
p9 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p10 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p11 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p12 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()

grid.arrange(p9, p10, p11, p12, ncol = 2)

```
**Based on the data:**
* Ascore: People with more 'agreeableness' show less usage - though 'agreeableness' is a little ambiguous. Agreeable to what - agreeable to peer pressure or agreeable to data that could tell them smoking is bad? A little more clarity could support this variable a bit better. 
* Cscore: People who are more conscientious tend to use nicotine less
* Impulsiveness: Impulsive people show higher amounts of usage than those who are not impulsive by nature
* SS: Not 100% sure what 'sensation seeing' is (as the guide describes). A more reasonable description would be 'sensation seeking' which is a psychological scale used and would make sense for this context. Based on that understanding, it makes sense why more sensation seekers use more nicotine. 

### Task 4: Create a random forest model on the training set to predict Nicotine using all of the variables in the dataset. You 5-fold, k-fold cross-validation (random number seed of 123 for the folds). Allow R to select mtryvalues between 2 and 8 and min_n values between 5 and 20. Use 10 levels in your “grid_regular” function.Set a random number seed of 123 for the tune_grid function. Use 100 trees.

  
```{r}

# Set up our folds for cross-validation
set.seed(123)
rf_folds = vfold_cv(train, v = 5)

```


```{r}

# Random forest with an R-defined tuning grid
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% 
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid 
)

```

```{r}

# Viewing Parameter Performance (Code credit https://juliasilge.com/blog/sf-trees-random-tuning/)
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")

```
```{r}

# Alternate View of Data
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")

```
### Task 5: Use the best mtry and min_n values from Task 4 to finalize the workflow and fit the model totraining set. Examine variable importance. What variables are most important in this model? (Hint: Referback to the dataset’s webpage if you need clarification as to meaning of any variables).
```{r}
# Finding the best of the values and saving them for use
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```


```{r}

# Saving Final Fit 
final_rf_fit = fit(final_rf, train)

```

```{r}
# Final fit Plot (Check out variable importance)
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")

```
**Based on the Data**
It appears the the number one important variable is SS. Going backward, the UK specifically and Oscore would be the next more important. 

### Task 6: How does the model perform on the training and testing sets?
```{r}

# Predictions on Training Set
trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)

```

```{r}

# Confusion Matrix on Training Set

confusionMatrix(trainpredrf$.pred_class, train$Nicotine, positive = "Yes")

```
```{r}

# Predictions on Test Data
testpredrf = predict(final_rf_fit, test)
head(testpredrf)


```
```{r}

# Confusion Matrix on Test
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")

```
*Based on the Data:*

The accuracy of the training set is 0.8861 and the testing set is 0.7325. Though we see a 0.1536 decrease in accuracy - it's still better than the Naive model of 0.6709. 

### Task 7: Comment on how this model might be used in the “real-world.” Would you recommend this modelfor real-world use? What if any concerns would you have about using the model?

In the real world, there are a couple of ways this data could be used. Considering the subject matter, a less ethical application (in my opinion) would be for tobacco companies to use this data to determine a specific demographic to market products - in order to harmonize the balance between sales and marketing efforts. On the flip side, health care professionals can use this data to consider the various implications of race and other demographics to better treat patients who may experience adverse effects of too much nicotine; and as a result, what medicines may be good or bad in the treatment of illnesses as a result. 

